\section{Mathematical models allow functional analysis of biological systems}

A variety of mechanistic mathematical modelling approaches has been used to analyze and elucidate function of cellular systems and viral infections.

For example, a process through which molecular motors may exert force onto influenza virus capsid layer is called tug-of-war. Tug-of-war involves velocity and force balance between all involved parties. Previously it has been used to describe two types of microtubule motors to model endosomal transport \cite{muller2008tug}, and to prove that competitive and cooperative modes of molecular motor transport are not mutually exclusive.

A stochastic model of bidirectional microtubular transport \cite{gazzola2009stochastic} is used to describe transport of adenovirus, and allows to predict the number of recruited molecular motors and available motor binding sites on the virus.

In another example, a biophysical model with  Langevin-type approximations  describes conformational changes of endosomolytic proteins in nonenveloped viruses \cite{lagache2012modeling}. This model is used to predict escape time based on the size of the endosome and the number of viral particles inside of the endosome.

A multiscale structured model of influenza infection \cite{heldt2013multiscale} predicts that drug interventions primarily targeting viral entry lead to simple delay of infection, while interventions targeting viral production allow to achieve reduction in viral titer.

These examples highlight that mathematical models coupled with existing knowledge of biological systems allow to make advanced predictions about underlying processes in these systems.