%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgements}{acknowledgements}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgements}

\def\thanks#1{%
\begingroup
\leftskip1em
\noindent #1
\par
\endgroup
}

I would like to thank Prof. J\"org Stelling for giving me this opportunity to work on this thesis, while following my interests, for supporting me whenever I asked for help, for his encouragements
and enthusiasm, and his creativity in solving problems. Thank you, Jörg, for all the support
throughout the years, all the valuable guidance and scientific freedom.
I would like to thank Adrian Egli for his essential support in realizing the projects presented in
this thesis and for always having an open ear for problems. Thank you, Adrian, for engaging
in the adventure of this interdisciplinary PhD project – having me as one of the first PhD
students circulating between the D-BSSE and the Department of Biomedicine, thank you for
introducing me to Immunology and for giving me the chance to learn so much over the last
years.
I thank my closest collaborator Yaseen Syedbasha for his great work, his tremendous help and
kindness, and for teaching me all the Tips and Tricks in the lab – together, we went through ups
and downs that are so characteristic to the research process – it was great fun to work together,
to discuss results that faced us sometimes as mysterious puzzles, to design new experiments,
shape projects, and support each other.
I would like to thank Michael Kaltenbach for introducing me to categorical regression, his
encouragements and mentorship throughout the years, his immense help in understanding
problems and shaping projects, his ability to shed light on difficult aspects, and countless
discussions that were of great importance to this thesis.
I thank Claudia Stühler for always having an open ear for immunological questions and for
helping me to orient through the wealth of immunological literature. I thank Fabian Rudolf
for asking the right questions, his encouragements, and for helping me with establishing the
avidity measurements. I thank Lukas Kaufmann and Dominik Vogt for their great work with
managing and measuring patient samples, which were of such importance for this thesis. I
thank Deborah Marks for inviting me to visit her lab and for valuable discussion. I thank her
and Chris Sander for their warm hospitality and the inspiring time I had during my visit.
I thank the past and current members of the Computational Systems Biology group for all the
valuable discussions, emotional support, great work environment and friendships. I also thank
the past and current members of the Applied Microbiology Research group for always being
extremely welcoming, supportive and also for providing a great work environment.
Finally, I would like to thank my highschool teacher Hermann Henne, who enabled me to visit
a summer school on Simulations as the third pillar of science, which sparked my fascination
for modelling living systems and greatly influenced my path. I would like to acknowledge the
support of the Studienstiftung during my undergraduate studies, which enabled me to study at
three different universities and follow my interests. And I would like to thank SystemsX for
supporting interdisciplinary research and organizing wonderful retreats for PhD students.

\endgroup
